Paper Analysis

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
added mu_beta sim plot. Also cleaned beta_sim plots chunk

# Winner effects on single dataset

# #the simulated dataset and corresponding fit to plot
# sim_dataset_num = 3
# 
# #global effects
# facet_labels = as_labeller(c(
#   mu_beta = "mu[beta]",
#   gamma1 = "gamma[1]",
#   gamma2 = "gamma[2]",
#   sigma_1 = "sigma[1]",
#   `tau[1]` = "tau[1]",
#   `tau[2]` = "tau[2]",
#   sigma_g1 = "sigma[g[1]]",
#   sigma_g2 = "sigma[g[2]]"
# ), label_parsed)
# 
# global_plot = sim_fit_list[[sim_dataset_num]]$draws(c("mu_beta", "gamma1", "gamma2", "sigma_1", "tau[1]", "tau[2]", "sigma_g1", "sigma_g2")) |> 
#   as_draws_df() |> 
#   pivot_longer(cols = "mu_beta":"sigma_g2") |> 
#   ggplot(aes(value)) +
#   geom_histogram() +
#    facet_wrap(~name, ncol = 4, scales = "free",
#              labeller = labeller(name = facet_labels)) +
#   labs(x = element_blank(), y = element_blank()) +
#   theme_4col() +
#   scale_x_continuous(breaks = breaks_pretty(n = 2)) 
# global_plot
# 
# 
# #experiential effects
# mod_pars = sim_fit_list[[sim_dataset_num]]$draws() |>
#     as_draws_df() |>
#     select(-any_of(c("lp__", ".chain", ".iteration", ".draw"))) |>
#     pivot_longer(cols = everything()) |>
#     filter(startsWith(name, "beta[2")) |>
#     mutate(id = stringr::str_extract(name, pattern = "\\d+]"),
#            id = stringr::str_replace(id, "\\]", ""), 
#            id = as.numeric(id))
# 
# #order of min to max estiamted betas - makes plot look nicer
# beta_order = mod_pars %>% group_by(id) %>% summarise(mean = mean(value)) %>% arrange(mean) %>% pull(id)
# 
# #true experiential effects
# true_betas = sim_datasets_list[[sim_dataset_num]] %>% 
#   group_by(player_id) %>% 
#   summarise(true_beta = unique(beta)) %>%
#   rename(id = player_id)
# 
# #reordering
# true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
#   mutate(ymin = as.numeric(factor(row_number())) - 0.2,
#          ymax = as.numeric(factor(row_number())) + 0.2)
# 
# #plot
# winner_plot = mod_pars |> 
#   ggplot(aes(y = reorder(id, value, FUN = median), x = value)) +
#   stat_histinterval() +
#   geom_segment(data = true_betas,
#                aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
#                color = "orange", linetype = "dashed", linewidth = 2) +
#   labs(x = "Estimated Experiential Effect", y = "Simulated Player ID") +
#   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
#   theme_single_y()
# winner_plot

@@

draws_list = lapply(sim_fit_list, function(x) {
  x$draws() |>
    as_draws_df() |>
    select(-any_of(c("lp__", ".chain", ".iteration", ".draw"))) |>
    pivot_longer(cols = everything()) |>
    filter(startsWith(name, "beta[2")) |>
    mutate(id = stringr::str_extract(name, pattern = "\\d+]"),
           id = stringr::str_replace(id, "\\]", ""), 
           id = as.numeric(id))
})

#combing fit results by mu_beta size
small_effect_draws = draws_list[1:3] %>% bind_rows(.id = "column_label") %>% 
  mutate(num_games = case_when(endsWith(column_label, "1") ~ 1000,
                               endsWith(column_label, "2") ~ 5000,
                               endsWith(column_label, "3") ~ 20000),
         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here

med_effect_draws = draws_list[4:6] %>% bind_rows(.id = "column_label") %>% 
  mutate(num_games = case_when(endsWith(column_label, "4") ~ 1000,
                               endsWith(column_label, "5") ~ 5000,
                               endsWith(column_label, "6") ~ 20000),
         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here

big_effect_draws = draws_list[7:9] %>% bind_rows(.id = "column_label") %>% 
  mutate(num_games = case_when(endsWith(column_label, "7") ~ 1000,
                               endsWith(column_label, "8") ~ 5000,
                               endsWith(column_label, "9") ~ 20000),
         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here

## small mu_beta plot ##

#true experiential effects
true_betas = sim_datasets_list[[1]] %>% 
  group_by(player_id) %>% 
  summarise(true_beta = unique(beta)) %>%
  rename(id = player_id)

#reordering
true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
         ymax = as.numeric(factor(row_number())) + 0.3)

#plot
small_effect_plot = small_effect_draws  |>
  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  #stat_dots(aes(colour = num_games), position = "dodgejust") +
  geom_segment(data = true_betas,
               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
ggsave(plot = small_effect_plot, 
       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_1.png"), dpi = 1000,
       height = 6, width = 7)

@@

#true experiential effects
true_betas = sim_datasets_list[[4]] %>% 
  group_by(player_id) %>% 
  summarise(true_beta = unique(beta)) %>%
  rename(id = player_id)

#reordering
true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
         ymax = as.numeric(factor(row_number())) + 0.3)

#plot
med_effect_plot = med_effect_draws  |>
  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  #stat_dots(aes(colour = num_games), position = "dodgejust") +
  geom_segment(data = true_betas,
               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
ggsave(plot = med_effect_plot, 
       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_2.png"), dpi = 1000,
       height = 6, width = 7)

@@

#true experiential effects
true_betas = sim_datasets_list[[7]] %>% 
  group_by(player_id) %>% 
  summarise(true_beta = unique(beta)) %>%
  rename(id = player_id)

#reordering
true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
         ymax = as.numeric(factor(row_number())) + 0.3)

#plot
big_effect_plot = big_effect_draws  |>
  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  #stat_dots(aes(colour = num_games), position = "dodgejust") +
  geom_segment(data = true_betas,
               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
ggsave(plot = big_effect_plot, 
       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_3.png"), dpi = 1000,
       height = 6, width = 7)
```

>>>>

@@

beta_draws_list = lapply(sim_fit_list, function(x) {
  x$draws() |>
    as_draws_df() |>
    select(-any_of(c("lp__", ".chain", ".iteration", ".draw"))) |>
    pivot_longer(cols = everything()) |>
    filter(startsWith(name, "beta[2")) |>
    mutate(player_id = stringr::str_extract(name, pattern = "\\d+]"),
           player_id = stringr::str_replace(player_id, "\\]", ""), 
           player_id = as.numeric(player_id))
})

#true experiential effects
true_small_betas = sim_datasets_list[[1]]$beta %>% unique()
true_med_betas = sim_datasets_list[[4]]$beta %>% unique()
true_big_betas = sim_datasets_list[[7]]$beta %>% unique()

beta_draws = beta_draws_list %>% bind_rows(.id = "dataset_id") %>%
  rename(draw = value) %>%
  mutate(dataset_id = substr(dataset_id, nchar(dataset_id), nchar(dataset_id)),
         dataset_id = as.numeric(dataset_id),
         true_beta = case_when(dataset_id %in% 1:3 ~ true_small_betas[player_id],
                               dataset_id %in% 4:6 ~ true_med_betas[player_id],
                               dataset_id %in% 7:9 ~ true_big_betas[player_id]),
         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
                               dataset_id %in% c(2, 5, 8) ~ 5000,
                               dataset_id %in% c(3, 6, 9) ~ 20000),
         num_games = factor(num_games, levels = c("1000", "5000", "20000")))

#for the geom_segment below
true_betas_plot = data.frame(true_small_val = true_small_betas, 
                             true_med_val = true_med_betas, 
                             true_big_val = true_big_betas, 
                             player_id = 1:length(true_small_betas)) %>%
  arrange(true_small_val) %>% #all orders are the same
  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
         ymax = as.numeric(factor(row_number())) + 0.3)

small_effect_plot = beta_draws %>% filter(dataset_id %in% 1:3) %>%
  ggplot(aes(y = reorder(player_id, true_beta), x = draw, colour = num_games)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  geom_segment(data = true_betas_plot,
               aes(x = true_small_val, xend = true_small_val, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
# ggsave(plot = small_effect_plot,
#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_1.png"), dpi = 1000,
#        height = 6, width = 7)

@@

med_effect_plot = beta_draws %>% filter(dataset_id %in% 4:6) %>%
  ggplot(aes(y = reorder(player_id, true_beta), x = draw, colour = num_games)) +
  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  #stat_dots(aes(colour = num_games), position = "dodgejust") +
  geom_segment(data = true_betas_plot,
               aes(x = true_med_val, xend = true_med_val, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
# ggsave(plot = med_effect_plot,
#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_2.png"), dpi = 1000,
#        height = 6, width = 7)

@@

#plot
big_effect_plot = beta_draws %>% filter(dataset_id %in% 7:9) %>%
  ggplot(aes(y = reorder(player_id, true_beta), x = draw, colour = num_games)) +
  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  #stat_dots(aes(colour = num_games), position = "dodgejust") +
  geom_segment(data = true_betas_plot,
               aes(x = true_big_val, xend = true_big_val, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
# ggsave(plot = big_effect_plot,
#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_3.png"), dpi = 1000,
#        height = 6, width = 7)
```

### mu_beta simulation comparison

```{r}
## same above plot but now mu_beta
mu_beta_draws_list = lapply(sim_fit_list, function(x) {
  x$draws("mu_beta") |>
  as_draws_df()
})

#true values of mu_beta
true_mu_betas = sim_datasets_list %>% bind_rows() %>% pull("mu_beta") %>% unique()

#getting the draws
mu_beta_draws = mu_beta_draws_list %>% bind_rows(.id = "dataset_id") %>%
  rename(draw = mu_beta) %>%
  mutate(dataset_id = substr(dataset_id, nchar(dataset_id), nchar(dataset_id)),
         dataset_id = as.numeric(dataset_id),
         true_mu_beta = case_when(dataset_id %in% 1:3 ~ true_mu_betas[1],
                                  dataset_id %in% 4:6 ~ true_mu_betas[2],
                                  dataset_id %in% 7:9 ~ true_mu_betas[3]),
         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
                               dataset_id %in% c(2, 5, 8) ~ 5000,
                               dataset_id %in% c(3, 6, 9) ~ 20000),
         true_mu_beta = factor(true_mu_beta, levels = c(as.character(true_mu_betas))),
         num_games = factor(num_games, levels = c("1000", "5000", "20000")))

#for the geom_segment below
true_mu_betas_plotting = data.frame(true_val = true_mu_betas) %>%
  mutate(ymin = as.numeric(factor(true_val)) - 0.25,
         ymax = as.numeric(factor(true_val)) + 0.25)

#plotting
mu_beta_plot = mu_beta_draws %>% ggplot(aes(y = true_mu_beta, x = draw, colour = num_games)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  geom_segment(data = true_mu_betas_plotting,
               aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

#save
# ggsave(plot = mu_beta_plot,
#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_mu_beta.png"), dpi = 1000,
#        height = 6, width = 7)
```
